Phylogenetics and the human microbiome

### 2013-04-01 sbreview.tex
something like an outline

The human microbiome is the collection of cells that live inside and on humans.
The Human Microbiome Project (HMP) and the Metagenomics of the Human Intestinal Tract (MetaHIT) consortium have advanced human microboime research recently by doing lots of sequencing, as well as bioinformatics and statistical development.
In this paper I review the impact of phylogenetics and tree-thinking on the methods used in the analysis of this data, as well as describing current challenges for phylogenetics coming from this type of work.

@@

% The HMP planned to comprehensively characterize the taxonomic
% composition of the microbiome by averaging 5,000 454 FLX 16S rRNA gene
% sequences from all 300 subjects, 18 body sites, and multiple time
% points. This design, combined with more than a 1,000-fold increase in
% sequencing throughput over the course of the HMP, forced the
% consortium to develop novel tools for processing large 16S rRNA gene
% datasets, tackling issues specific to 454 sequence data quality, and
% addressing novel biological questions that were previously
% inaccessible due to limited sample sizes.

Now we will embark on our journey.

There have been hundreds of papers concerning the human microbiome. It
is not our desire to survey all of these. Rather, our idea is to
highlight areas that may be of particular interest for the
phylogenetics community by potentially having wider applications or by
being areas for future exploration.

Please note that many of the methods and issues described in this paper will be equally useful for microbial ecology in general, but by narrowing our scope we were able to keep our paper within reasonable size limits.

Antibiotics.
\cite{dethlefsen2008pervasive,dethlefsen2011incomplete,jakobsson2010short,jernberg2007long}

@@

\subsection{phylogenetics and taxonomy}

@@

However, there have been

There are projects that use phylogeny to revise taxonomy, short of PhyloCode.
GreenGenes.
The reason people are attached to taxonomy is that there is a lot of information using taxonomic labels.

@@

With the advent of high throughput sequencing, there has been a lot of interest concerning which variable region to use.

@@

@@

@@

@@

@@

\subsection{The challenge that lies ahead: functional genes and phylogenomics}

@@

@@

\bibliographystyle{plainnat}
\bibliography{sbreview}
\end{document}

>>>>

The human microbiome is the collection of cells that live inside and on humans.
The Human Microbiome Project (HMP) and the Metagenomics of the Human Intestinal Tract (MetaHIT) consortia have advanced human microboime research by generating thousands of 16s surveys and terabases of metagenomic data, as well as funding bioinformatics and statistical development.
In this paper I review the impact of phylogenetics and tree-thinking on the methods used in the analysis of this data, as well as describing current challenges for phylogenetics coming from this type of work.

@@

Starter paragraph about the scope of human microbiome research.

We will begin by introducing the results of recent research.
There have been hundreds of papers concerning the human microbiome.
It is not our desire to survey all of these.
Rather, our idea is to highlight areas that may be of particular interest for the phylogenetics community by potentially having wider applications or by being areas for future exploration.
Please note that many of the methods and issues described in this paper will be equally useful for microbial ecology in general, but by narrowing our scope we were able to keep our paper within reasonable size limits.

Major finding: dynamic in terms of representation but diverse in terms of function.

Major finding: antibiotics.
\cite{dethlefsen2008pervasive,dethlefsen2011incomplete,jakobsson2010short,jernberg2007long}

Enterotypes discussion.

@@

\subsection{Phylogenetics and taxonomy}

@@

However, there have been a number projects that use phylogeny to revise taxonomy, short of PhyloCode.
GreenGenes.
tax2tree.
Bergey's taxonomy is supposed to be.
SILVA.
The reason people are attached to taxonomy is that there is a lot of information using taxonomic labels.

@@

Tree-based classification.

Curation of HM databases using phylogenetics?

@@

Alpha diversity.

@@

Beta diversity.

@@

\subsection{Inference of horizontal gene transfer}

@@

ARB software.
Phylogenetic placement.

@@

Non-overlapping reads.
Assemblies.

Functional genes and phylogenomics.

HGT.

@@

Field is still not yet mature.

@@

\bibliographystyle{plainnat}
\bibliography{sbreview}

\end{document}
